---
title: "Rebecca Van Hoeck"
author: "Becca Van Hoeck"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lunar)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

```

## Script to build models and evaluate model fit to identify environmental drivers of atlantic cod vocalizations

Load data - all data wrangling and exploratory visualizations were done in historical_cod_data.RMD
```{r}

## Tidy data with grunt presence, rate, and associated environmental variables
# note that there are many times missing temperatures. Possibly not able to use this variable because grunt data is so sparse. I don't want to remove some cod data just to have complete cases with temperature
cod_data = read.csv("data/cod_allregions07-21.csv", header = TRUE)
cod_data = cod_data[,-1] # remove row index

```

## Creating sin and cos variables for natural cycles

```{r}

# Hour of the day
cod_data$shour = sin(2*pi*(cod_data$hour/24)) # this matches
cod_data$chour = cos(2*pi*(cod_data$hour/24)) 

# Julian Day
cod_data$sday = sin(2*pi*(cod_data$J/365))
cod_data$cday = cos(2*pi*(cod_data$J/365))
  
# Lunar cycle  
cod_data$sLunar = sin(cod_data$lunarphase) 
cod_data$cLunar = cos(cod_data$lunarphase)

#Semi lunar cycle
# unsure how to calculate, asked colleague

  

```

# Mixture and hurdle models
Caiger et al 2020 (MEPS) Global Model: Ten years of data from Mass Bay

presence or rate ~ year + site + hour + depth + lunar + semiLunar + JulianDay + S:Julian Day + random(week)

- hour, lunar, semilunar, julian day were all cyclical
- random effect of week was to account for serial autocorrelation
- Used model averaged predictions, weighted by AIC, with all models with deltaAIC <5. 

Presence model: Top 4 models averaged. All included Y, S, H, D, J, rW, S:J. One included L1, one omitted all Lunar, one included both L1 & L2, one included only L1. 

Rate model: model evaluated predictors for the rate zero-inf terms separately. Two models were averaged, rate terms were the same and included: Y, S, H, L1, L2, J, rW, S:J. Zero inflated were the same as rate, except one model excluded L2. 

```{r}
# following ENEC563 Lab 15 for Mixture and Hurdle Models

# starting with the global model
# does not include region yet, because of singularity with site
# also does not include random effect of week. 

library(pscl)
zip1 = zeroinfl(n_grunts ~ SpawnSeason + site + shour + chour + sLunar + cLunar + sday + cday + site:J |
                  SpawnSeason + site + shour + chour + sLunar + cLunar + sday + cday + site:J, 
                dist = "poisson", link = "logit", data = cod_data)
summary(zip1)

#this was computationally singular? 
Nb1 = zeroinfl(n_grunts ~ SpawnSeason + site + shour + chour + sLunar + cLunar + sday + cday + site:J |
                  SpawnSeason + site + shour + chour + sLunar + cLunar + sday + cday + site:J, 
                dist = "negbin", link = "logit", data = cod_data)
summary(Nb1)

# but the likelihood is much higher for the Neg Bin that only gave coefficient estimates
library(lmtest)
lrtest(zip1, Nb1)


```

# Modeling presence and rate separately

# Presence with Cox Ledge and Mass Bay Data

```{r}
# singularity for region when site is included
codMod = glm(presence ~ SpawnSeason + site + region + shour + chour + sLunar + cLunar,
                 data = cod_data, family = binomial(link = "logit"))
summary(codMod)

codMod_nosite = glm(presence ~ SpawnSeason + region + shour + chour + sLunar + cLunar,
                 data = cod_data, family = binomial(link = "logit"))
summary(codMod_nosite)
drop1(codMod_nosite)
cod_data$P_resid = residuals(codMod_nosite, type = "pearson")
cod_data$d_resid = residuals(codMod_nosite, type = "deviance")

sum(cod_data$P_resid)/df.residual(codMod_nosite)
# does this mean that there is a significant difference in presence, or significant difference in covariates?

```

## Model Grunt Rate

```{r}
# Rate Model
library(gamlss)
# Full model - need to add semi-lunar cycle
RateMod1 = glm(n_grunts ~ SpawnSeason + region + shour + chour + sLunar + cLunar, 
               data = cod_data, poisson(link = "log"))
summary(RateMod1)

# same model but negative binomial
RateModNB = glm.nb(n_grunts ~ SpawnSeason + region + shour + chour + sLunar + cLunar, data = cod_data)
summary(RateModNB)

# same model, but quadratic error term on negative binomial
# why remove nas in this package only?
# RateModNBII = gamlss(n_grunts ~ SpawnSeason + region + shour + chour + sLunar + cLunar,
#                      data = na.omit(cod_data), family = NBII)
# summary(RateModNBII)

#AIC(RateMod1, RateModNB, RateModNBII, RateModNBII_2)


#ratePredict = predict(RateModNBII)

#plot(ratePredict)
```

